Advantages of mixed unitary operators for quantum information processing

### 2018-12-26 decorrelation/mixed_unitaries.tex
more plots, cleaning up convex code and notebooks. experimental and numerical section should be in 80% good shape.

The top row of all trace-preserving (TP) maps is fixed to $\{1,0,0,0,\cdots\}$.  The first column, $\vec{m}$, describes any deviations from unitality, such as could arise from amplitude damping. If the error channel is unitary, then the error is coherent, and the submatrix $R$ is perfectly antisymmetric, corresponding to a rotation of the generalized Bloch vector. If  $R$ is diagonal, then the error channel is Pauli stochastic, with each entry corresponding to the probability that the associated Pauli error occurs in each application of the gate. \note{Be more clear about what "associated" means here}.  If R is symmetric but not diagonal, then the channel is still stochastic, but the random errors consist of correlated Pauli operators (such as $X+Y$). For a single qubit, this describes everything, but the situation can be slightly more complicated for more qubits.

@@

The \emph{size} of an error in quantum gates may  be quantified in a number of ways. Two of the most metrics are the average gate fidelity, $\mathcal{F}$, and the diamond norm, $\vert\vert\cdot\vert\vert_\diamond$. These may be represented in terms of the error maps as
\begin{equation}
	\vert\vert \error \vert\vert_\diamond = \sup_\rho \vert \vert (I\otimes I)(\rho) - (\error \otimes I)(\rho) \vert\vert_1
\end{equation}
\begin{equation}
	\mathcal{F}(\error) = ...
\end{equation}
The diamond norm is generally linear in the over-rotation angle of a quantum operation. Also definition of average gate infidelity. The AGI is generally quadratic in the over-rotation angle of a quantum operation.

@@

Because we our trying to generate high-order RBCs, we want to generate controls that perform well even in the presence of drift. To do this, we modified the gradient in GRAPE to instead be:
\begin{align}\label{quadrature}
\frac{\partial \tilde J(U)}{\partial u_{ij}} =
\int p(\vec{\delta})\frac{\partial J(U(\vec{\delta}))}{\partial u_{ij}} d\vec{\delta}
\end{align}
with $p(\vec{\delta})$ Gaussian distributed, as has been done in previous works \cite{Goerz2014} to ensure that the optimal control results are robust over a wide range of errors. To make this averaging tractable, we approximate this integral using Gaussian quadrature, approximating the cost function as a low order polynomial\cite{abramowiz1972handbook} (in particular degree 5). Note that despite updating the gradient, we did not change the cost function. The termination condition for the gradient ascent loop is still just that the controls perform well in the absence of drift. This is not a fundamental requirement of the routine, but rather was done to make finding solutions with $J(U) < 1 - \epsilon$  less computationally taxing.

% ==============================================================================
% Section: Numerical Results
% ==============================================================================
\section{Numerical Results}
\label{sec:numerical_results}
In the following numerical results, we explore using the methods in Section \ref{sec:mixed_unitary_processes} to build MUPs. We consider the following model for a single tunable qubit: 
\begin{equation}\label{eq:1Qham}
  H(\delta, \epsilon, t) = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
We use the GRAPE algorithm as discussed in Section  \ref{sec:random_gate_synthesis} with N=25 steps and total evolution time of $\pi$ to generate 100 candidate controls, with the Gaussian-weighted gradient averaged over a standard deviation of $\sigma=.001$. We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as in superconducting qubit systems that share a common RF drive.

Solving the optimization problem defined in  \ref{sec:robustly_mixed} yields MUPs for $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$ shown in Figure \ref{fig:XMUP} and Figure \ref{fig:YMUP}.

These results demonstrate several properties that make MUPs useful and tractable to generate. First, naively generating a 0MUP results in nontrivial support on all the members of the control family. However, by rewriting the minimization to impose the sparsity constraint discussed in \ref{sec:robustly_mixed}, we can generate a 0MUP that uses just five of the controls. This shows that through adding constraints to our optimization routine, we can make the MUP practically useful, i.e. not requiring a prohibitive amount of waveform memory. When generating a 1MUP we impose the $\ell_1$ penalty discussed in \ref{sec:robustly_mixed}, so that the algorithm favors smaller error Hamiltonians. While the solution in this case was already sparse, enforcing the penalty allows for a smooth trade off between performance at the origin, and robustness to drift.

@@

In this example it was infeasible to use GRAPE to return non-trivial solutions. Instead we manually selected piecewise constant echoing sequences with 500 steps and total evolution time of $\frac{5\pi}{2}$. In particular, we considered $RX(\pi)$, $RX(-\pi)$, $RY(\pi)$ and $RY(-\pi)$ pulses used to decouple detuning errors as taking only one time step, to avoid significant errors arising from non-commutativity. The specific collection of bang-bang sequences \cite{bangbang} considered were all combinations of simultaneous $\pi$ pulses activated at multiples of $8$ steps from the first time step, and the same multiple of $8$ steps prior to the last time step. To give the control family a variety of RF errors, we added on uniformly distributed errors to each $\pi$ pulse, between $-.25$\% and $.25$\%.

In this example, we find much more modest improvements to performance, as shown in Figure \ref{fig:2MUP}. Indeed, there are now four free parameters to optimize over, and the uncontrolled entangling interaction means that there is little variation in the controls. Nonetheless, using a MUP improves performance by half of an order of magnitude at the origin, and up to an order of magnitude away from the origin, with the 1MUP performing just as well as the 0MUP, but also being more robust to drift. We believe that if we used a two qubit interaction that was controllable, such as tunable couplers, there would be room for more interesting controls and perhaps even better performance improvements.
% ==============================================================================
% Section: Experimental Results
% ==============================================================================
\section{Experimental Results}
\label{sec:experimental_results}
Here we present experimental results from implementing our routine on a fixed-frequency superconducting transmon qubit. In particular, we used qubit 8 on the Rigetti 19Q-Acorn chip, whose characterization can be found in \cite{1712.05771}. To implement a MUP on this qubit, four incorrectly calibrated Gaussian pulses were produced by scaling the pulseshape amplitude for a calibrated 10 sample 50ns $RX(\frac{\pi}{2})$ pulse by $106.4\%$,  $103.9\%$, $93.7\%$ and $91.2\%$.

@@

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{SQRTX.png}
  \caption{Numerical results comparing a 0MUP to a 1MUP for a single tunable qubit, for $RX(\frac{np.pi}{2})$. Shown with lower alpha values are example members of the control families, with diamond norms of around $10^{-2}$. The 0MUP can be seen to outperform the members of the control family by two orders of magnitude at the origin, and the 1MUP by an order of magnitude. However, the 1MUP can be seen to be flatter. In paritcular, for small detunings, the 1MUP performs better than the 0MUP.}
  \label{fig:1QMUP}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{SQRTY.png}
  \caption{Numerical results comparing a 0MUP to a 1MUP for a single tunable qubit, for $RY(\frac{\pi}{2})$. The results are quantitatively similar to those for $RX(\frac{\pi}{2})$, but in this case the 0MUP is worse at the origin. This can be improved by penalizing for the norms, or by simply trying to find a 1MUP, which both symmeterizes the performance at the origin and makes them better.}
  \label{fig:XMUP}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{2QRBC.png}
  \caption{Numerical results comparing a 0MUP to a 1MUP for a pair of tunable qubits, with a resonant exchange interaction. Shown with lower alpha values are example members of the control families. The 0MUP and 1MUP can be seen to outperform the members of the control family by half an order of magnitude at the origin. Away from the origin, the 1MUP can be seen to be flatter. Far away from the origin, the 1MUP outperforms members of the control famlies by almost an order of magnitude in diamond norm.}
  \label{fig:2MUP}
\end{figure*}
% ==============================================================================
% Section: Acknowledgements
% ==============================================================================
\section{Acknowledgements}
\label{sec:acknowledgements}
Sandia National Laboratories is a multimission laboratory managed and operated by National Technology and Engineering Solutions of Sandia, LLC, a wholly owned subsidiary of Honeywell International, Inc., for the U.S. Department of Energy's National Nuclear Security Administration under contract DE-NA0003525.
\bibliography{decorrelation.bib}
\end{document}

>>>>

The top row of all trace-preserving (TP) maps is fixed to $\{1,0,0,0,\cdots\}$.  The rest of the first column, $\vec{m}$, describes any deviations from unitality, such as could arise from amplitude damping. If the error channel is unitary, then the error is coherent, and the submatrix $R$ is perfectly antisymmetric, corresponding to a rotation of the generalized Bloch vector. If  $R$ is diagonal, then the error channel is Pauli stochastic, with each entry corresponding to the probability that the associated Pauli error occurs in each application of the gate. \note{Be more clear about what "associated" means here}.  If R is symmetric but not diagonal, then the channel is still stochastic, but the random errors consist of correlated Pauli operators (such as $X+Y$). For a single qubit, this describes everything, but the situation can be slightly more complicated for more qubits.

@@

The \emph{size} of an error in quantum gates may  be quantified in a number of ways. Two of the most common metrics are the average gate fidelity, $\mathcal{F}$, and the diamond norm, $\vert\vert\cdot\vert\vert_\diamond$. These may be represented in terms of the error maps as
\begin{equation}
	\vert\vert \error \vert\vert_\diamond = \sup_\rho \vert \vert (I\otimes I)(\rho) - (\error \otimes I)(\rho) \vert\vert_1
\end{equation}
\begin{equation}
	\mathcal{F}(\error) = ...
\end{equation}
The diamond norm is generally linear in the over-rotation angle of a quantum operation. Also definition of average gate infidelity. The AGI is generally quadratic in the over-rotation angle of a quantum operation.

@@

Because we our trying to generate high-order RBCs, we want to generate controls that perform well even in the presence of drift. To do this, we modified the gradient in GRAPE to instead be:
\begin{align}\label{eq:quadrature}
\frac{\partial \tilde J(U)}{\partial u_{ij}} =
\int p(\vec{\delta})\frac{\partial J(U(\vec{\delta}))}{\partial u_{ij}} d\vec{\delta}
\end{align}
with $p(\vec{\delta})$ Gaussian distributed, as has been done in previous works \cite{Goerz2014} to ensure that the optimal control results are robust over a wide range of errors. To make this averaging tractable, we approximate this integral using Gaussian quadrature, approximating the cost function as a low order polynomial\cite{abramowiz1972handbook} (in particular degree 5). Note that despite updating the gradient, we did not change the cost function. The termination condition for the gradient ascent loop is still just that the controls perform well in the absence of drift. This is not a fundamental requirement of the routine, but rather was done to make finding solutions with $J(U) < 1 - \epsilon$  less computationally taxing.

% ==============================================================================
% Section: Numerical Results
% ==============================================================================
\section{Numerical Results}
\label{sec:numerical_results}
In the following numerical results, we explore using the methods in Section \ref{sec:mixed_unitary_processes} to build MUPs. We consider the following model for a single tunable qubit: 
\begin{equation}\label{eq:1Qham}
  H(\delta, \epsilon, t) = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
We use the GRAPE algorithm as discussed in Section  \ref{sec:random_gate_synthesis} with N=25 steps and total evolution time of $\pi$ to generate 100 candidate controls, with a standard deviation of $\sigma=.001$ for the distribution in Equation \ref{eq:quadrature}. We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as is the case in systems that implement RZ rotations with phase shifts of the control signal. Solving the optimization problem defined in Section \ref{sec:robustly_mixed} yields similar MUPs for $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$, with the results for $RY(\frac{\pi}{2})$ shown in Figure \ref{fig:YMUP}. These results demonstrate several properties that make MUPs both useful and tractable.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{SQRTY.png}
  \caption{Numerical results comparing a 0MUP to a 1MUP for a single tunable qubit, for $RY(\frac{\pi}{2})$. The results are quantitatively similar to those for $RX(\frac{\pi}{2})$, but in this case the 0MUP is worse at the origin. This can be improved by penalizing for the norms, or by simply trying to find a 1MUP, which both symmeterizes the performance at the origin and makes them better.}
  \label{fig:YMUP}
\end{figure*}

%\begin{figure*}
%  \centering
%  \includegraphics[width=\textwidth]{SQRTX.png}
%  \caption{Numerical results comparing a 0MUP to a 1MUP for a single tunable qubit, for $RX(\frac{np.pi}{2})$. Shown with lower alpha values are example members of the control families, with diamond norms of around $10^{-2}$. The 0MUP can be seen to outperform the members of the control family by two orders of magnitude at the origin, and the 1MUP by an order of magnitude. However, the 1MUP can be seen to be flatter. In paritcular, for small detunings, the 1MUP performs better than the 0MUP.}
%  \label{fig:XMUP}
%\end{figure*}

First, naively generating a 0MUP results in nontrivial support on all the members of the control family. However, by rewriting the minimization to impose the sparsity constraint discussed in Section \ref{sec:robustly_mixed}, we can generate a 0MUP that uses just five of the controls. This shows that through adding constraints to our optimization routine, we can make the MUP practically useful. In both cases we impose the same $\ell_2$ penalty as described in Section \ref{sec:robustly_mixed}, so that the algorithm preferentially selects controls with smaller errors. Imposing this constraint allows us to trade off flatness at the origin for performance.

@@

In this example it was infeasible to use GRAPE to return non-trivial solutions. Instead we manually selected piecewise constant echoing sequences with 500 steps and total evolution time of $\frac{5\pi}{2}$. In particular, we considered $RX(\pi)$, $RX(-\pi)$, $RY(\pi)$ and $RY(-\pi)$ bang-bang sequences \cite{bangbang}, consisting of all combinations of simultaneous $\pi$ pulses activated at multiples of $8$ steps from the beginning of the controls, and the same multiple of $8$ steps prior to the end of the controls. An example of such a control is given in Figure \ref{fig:bangbang}. To give the control family a variety of RF errors, we added on uniformly distributed errors to each $\pi$ pulse, between $-.25$\% and $.25$\%.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{bangbang.png}
  \caption{Example bang-bang sequence, using both qubits' $\sigma_x$ controls. Not shown are the $\sigma_y$ controls, and the uncontrolled resonant exchange interaction.}
  \label{fig:bangbang}
\end{figure}

In this example, we find more modest improvements to performance, as shown in Figure \ref{fig:2MUP}. There are now four free parameters to optimize over, and the uncontrolled entangling interaction means that there is little variation in the controls. Nonetheless, using a MUP improves performance by half of an order of magnitude at the origin, and up to an order of magnitude away from the origin, with the 1MUP performing just as well as the 0MUP, but also being more robust to drift.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{2QRBC.png}
  \caption{Numerical results comparing a 0MUP to a 1MUP for a pair of tunable qubits, with a resonant exchange interaction. Shown with lower alpha values are example members of the control families. The 0MUP and 1MUP can be seen to outperform the members of the control family by half an order of magnitude at the origin. Away from the origin, the 1MUP can be seen to be flatter. Far away from the origin, the 1MUP outperforms members of the control famlies by almost an order of magnitude in diamond norm.}
  \label{fig:2MUP}
\end{figure*}

% ==============================================================================
% Section: Experimental Results
% ==============================================================================

\section{Experimental Results}
\label{sec:experimental_results}
Here we present experimental results from implementing our routine on a fixed-frequency superconducting transmon qubit. In particular, we used qubit 8 on the Rigetti 19Q-Acorn chip, whose characterization can be found in \cite{1712.05771}. To implement a MUP on this qubit, four incorrectly calibrated Gaussian pulses were produced by scaling the pulseshape amplitude for a calibrated 10 sample 50ns $RX(\frac{\pi}{2})$ pulse by $106.4\%$,  $103.9\%$, $93.7\%$ and $91.2\%$.

@@

% ==============================================================================
% Section: Acknowledgements
% ==============================================================================
\section{Acknowledgements}
\label{sec:acknowledgements}
Sandia National Laboratories is a multimission laboratory managed and operated by National Technology and Engineering Solutions of Sandia, LLC, a wholly owned subsidiary of Honeywell International, Inc., for the U.S. Department of Energy's National Nuclear Security Administration under contract DE-NA0003525.
\bibliography{decorrelation.bib}
\end{document}
